Visual Interrogation of Attention-Based Models: \\ Applications in Language Inference and Machine Comprehension

### 2018-05-29 EMNLP-demo-paper/NLPVis-demo-paper.tex
add code
added model description, very brief
add perturbation discussion
done with system
add author list

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{emnlp2018}
\usepackage{times}
\usepackage{latexsym}
\usepackage{url}
\usepackage{mathptmx}

%\aclfinalcopy % Uncomment this line for the final submission

@@

\usepackage{graphicx}
\graphicspath{{figures/}{pictures/}{figs/}{./}} % where to search for the images

\newcommand\BibTeX{B{\sc ib}\TeX}
\newcommand\confname{EMNLP 2018}
%\newcommand\conforg{SIGDAT}

@@

\author{Shusen Liu \\
  Lawrence Livermore National Laboratory\\
  {\tt liu42@llnl.gov} \\\And
  Tao Li,  Zhimin Li,  Vivek Srikumar, Valerio Pascucci \\
  School of Computing, University of Utah\\
    {\tt {tao.li, zhimin.li, vivek, pacucci}@utah.edu}
}

\date{}

\begin{document}
\maketitle

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{emnlp2018}
\usepackage{times}
\usepackage{latexsym}
\usepackage{url}
\usepackage{mathptmx} 
\usepackage{amsmath}

\aclfinalcopy % Uncomment this line for the final submission

@@

\usepackage{graphicx}
\graphicspath{{figures/}{pictures/}{figs/}{./}} % where to search for the images
\usepackage{xcolor}  % Coloured text etc.
\usepackage{listings}
%\usepackage{authblk}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
\lstdefinestyle{mystyle}{
    %backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\scriptsize\ttfamily\bfseries,
    %basicstyle=\footnotesize\ttfamily,
    breakatwhitespace=false,         
    %breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    %numbers=left,                    
    %numbersep=1pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=1
}

\lstset{style=mystyle}

\newcommand\BibTeX{B{\sc ib}\TeX}
\newcommand\confname{EMNLP 2018}

@@

\author{Shusen Liu$^{1}$, Tao Li$^{2}$,  Zhimin Li$^{3}$,  Vivek Srikumar$^{2}$, Valerio Pascucci$^{3}$, Peer-Timo Bremer$^{1}$ \\
 % {\tt liu42@llnl.gov} \\\And
  Lawrence Livermore National Laboratory$^{1}$\\
  School of Computing, University of Utah$^{2}$\\  
  SCI Institute, University of Utah$^{3}$\\
   % {\tt {tao.li, zhimin.li, vivek, pacucci}@utah.edu}
}

%\date{}
\begin{document}
\maketitle

%\vspace{-4mm}

### 2018-05-29 EMNLP-demo-paper/system.tex
add author list

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.9\linewidth]{sentence}
 \caption{
The interface for showing input sentences. The user can manually alter the words or apply automatic perturbation/paraphrasing of the inputs. In the ``perturbed'' dropdown menu, the blue words are the only not exist in the original sentence.
 }
\label{fig:sentence}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.8\linewidth]{prediction}
 \caption{
Summarize the prediction results of the perturbed input for the natural language inference model.
the prediction is encoded as a point in the barycentric coordinate system of the triangle, in which each vertex corresponds to one prediction label.
}
\label{fig:prediction}
\end{figure}

@@

\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=0.9\linewidth]{attentionPanels}
  \vspace{-3mm}
 \caption{
Attention visualization. In the graph attention view (a), a bipartite graph encoding is adopted, in which the edge thickness corresponds to the attention value. In the matrix attention view (b), the entries of $i^{th}$ row represent the probabilities of words in hypotheses align to the $i^{th}$ word in the premise.
The user can alter the attention values via the pop-up interface illustrated in (c).
We overlay the dependency tree ($a_1$) grammar structure to highlight important words and allow simplification of complex sentence based on the dependency tree.
%
For highly asymmetric attention relationship, we utilized a zoomable hierarchical visual representation (d).
}
\label{fig:attentionVis}
\end{figure*}

\subsection{Attention Visualization}
As illustrated in Figure~\ref{fig:attentionVis}(a)(b), the most widely adopted technique to bipartie graph (a), as well as the heatmap attention matrix representation (b). 
%
In the graph attention view (Fig.~\ref{fig:attentionVis}(a)), a bipartite graph encoding is adopted, in which the edge thickness corresponds to the attention value. The color of the rectangle text block encodes the sum of all edge values connected to it (darker shade of blues correspond to higher values).
%%%%%%%%
The graph view is suitable for highlighting the most dominant alignments. However, the edges may become cluttered if multiple attention values are high. The matrix attention view (Fig.~\ref{fig:attentionVis}(b)) resolve these issues, despite being more verbose and less efficient in highlighting the dominant alignments. 
We also enable the linkage between highlighted actions in both views (see Fig.~\ref{fig:attentionVis}(a)(b), one the most dominant attention is highlighted in orange).
%
To facilitate the perturbation of attention (see Figure~\ref{fig:pipeline}), as illustrated in Figure~\ref{fig:attentionVis}(c), we implement an interface in the matrix view for user to directly manipulate the attention value.

@@

However, when the one of the text sequence become signification longer, i.e., a full paragraph as in the machine comprehension task, the even the simplified sentence can not be meaningfully represented in the graph or matrix visual encoding. To address this visualization challenge, as illustrated in Figure~\ref{fig:attentionVis}(d), we introduce the hierarchical representation. Here, a pure graphical encoding (the color bars marked by \emph{att2}) is used to capture the aggregated alignment information of the whole paragraph, and the user can focus on localized attention information by selecting the a pixel bar that presents a single sentence. We can also link this attention representation with matrix, such that whenever a sentence is selected the local attention relationship is allow shown in the matrix view (see Figure~\ref{MCexample}(a)).

>>>>

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.9\linewidth]{sentence}
 \caption{
The interface for showing input sentences. The user can manually alter the words or apply automatic perturbation/paraphrasing of the inputs. In the ``perturbed'' dropdown menu, the blue words are the only not exist in the original sentence.
 }
 \vspace{-3mm}
\label{fig:sentence}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.8\linewidth]{prediction}
 \caption{
Summarize the prediction results of the perturbed input for the natural language inference model.
the prediction is encoded as a point in the barycentric coordinate system of the triangle, in which each vertex corresponds to one prediction label.
}
\vspace{-3mm}
\label{fig:prediction}
\end{figure}

@@

\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=0.9\linewidth]{attentionPanels}
  \vspace{-3mm}
 \caption{
Attention visualization. In the graph attention view (a), a bipartite graph encoding is adopted, in which the edge thickness corresponds to the attention value. In the matrix attention view (b), the entries of $i^{th}$ row represent the probabilities of words in hypotheses align to the $i^{th}$ word in the premise.
The user can alter the attention values via the pop-up interface illustrated in (c).
We overlay the dependency tree ($a_1$) grammar structure to highlight important words and allow simplification of complex sentence based on the dependency tree.
%
For highly asymmetric attention relationship, we utilized a zoomable hierarchical visual representation (d).
}
\vspace{-3mm}
\label{fig:attentionVis}
\end{figure*}

\subsection{Attention Visualization}
As illustrated in Figure~\ref{fig:attentionVis}(a)(b), the most widely adopted technique to bipartie graph (a), as well as the heatmap attention matrix representation (b). 
%
In the graph attention view (Fig.~\ref{fig:attentionVis}(a)), a bipartite graph encoding is adopted, in which the edge thickness corresponds to the attention value. The color of the rectangle text block encodes the sum of all edge values connected to it (darker shade of blues correspond to higher values).
%%%%%%%%
The graph view is suitable for highlighting the most dominant alignments. However, the edges may become cluttered if multiple attention values are high. The matrix attention view (Fig.~\ref{fig:attentionVis}(b)) resolve these issues, despite being more verbose and less efficient in highlighting the dominant alignments. 
We also enable the linkage between highlighted actions in both views (see Fig.~\ref{fig:attentionVis}(a)(b), one the most dominant attention is highlighted in orange).
%
To facilitate the perturbation of attention (see Figure~\ref{fig:modelPipeline}), as illustrated in Figure~\ref{fig:attentionVis}(c), we implement an interface in the matrix view for user to directly manipulate the attention value.

@@

However, when the one of the text sequence become signification longer, i.e., a full paragraph as in the machine comprehension task, the even the simplified sentence can not be meaningfully represented in the graph or matrix visual encoding. To address this visualization challenge, as illustrated in Figure~\ref{fig:attentionVis}(d), we introduce the hierarchical representation. Here, a pure graphical encoding (the color bars marked by \emph{att2}) is used to capture the aggregated alignment information of the whole paragraph, and the user can focus on localized attention information by selecting the a pixel bar that presents a single sentence. We can also link this attention representation with matrix, such that whenever a sentence is selected the local attention relationship is allow shown in the matrix view (see Figure~\ref{fig:MCexample}(a)).
